A Horrocks-type theorem for even orthogonal $K_2$

### 2020-04-16 dl-paper.tex
Fix misprints

Now set $\widetilde{G}_n = G(\rD_n, -) = \mathrm{Spin}(2n, -)$ and $\widetilde{P}_n = \mathrm{E}(\rD_n, -) = \mathrm{Epin}(2n, -)$.
Notice that for $n\geq 3$ the groups
\begin{align*}
   \KO_2(2n, R) &= \K_2^{G_n,P_n} = \Ker(\St(\rD_n, R) \to \mathrm{SO}(2n, R)), \\
 \K_2(\rD_n, R) &= \K_2^{\widetilde{G}_n,\widetilde{P}_n} = \Ker(\St(\rD_n, R) \to \mathrm{Spin}(2n, R))
\end{align*}
are related via the exact sequence
\[ \xymatrix{ 1 \ar[r] & \K_2(\rD_n, R) \ar[r] & \KO_2(2n, R) \ar[r] & \mu_2(R) \ar[r] & 1,} \]
which can be obtained from the exact sequence on p.~189 of~\cite{Ba74} using the snake lemma (cf. also with~\cite[Corollary~4.3.5]{Ba74}).
On the other hand, from~\cref{lem:lowerKgroups} it follows that $\KO_3(2n, R) \cong K_3^{\widetilde{G}_n, \widetilde{P}_n}$.

@@

\begin{lemma} \label{SR:obtuse} For $g \in j_+(K(A[X], M[X])) \cap P_\alpha(m) \cap P_\beta(m')$ and $a, b\in A$ the element \[ x_{\beta - \alpha}(b) \cdot S_\alpha(a, g) \cdot h_{\alpha}^{-1}(1+am)\cdot h_{\beta - \alpha}((1 + \epsilon abm')^{-1})\cdot x_{\beta - \alpha}(-b(1 + \epsilon abm')) \]
 belongs to $P_\alpha\left(\tfrac{m - \epsilon bm'}{1+am}\right)$, where $\epsilon = N_{\alpha, -\beta}$.
\end{lemma}
\begin{proof} Set $c = 1 + \epsilon a b m'$, $s_0 = \langle a, m \rangle$, $s_1 = \langle \epsilon am', -bc^{-1} \rangle$.
Notice that $1 - \epsilon abc^{-1}m' = c^{-1}$, therefore by~\eqref{eq:dennis-stein}
\begin{equation} \label{eq:computation-SR-obtuse} x_{\alpha-\beta}(\epsilon am') \cdot x_{\beta-\alpha}(-bc^{-1}) = x_{\beta-\alpha}(-b) \cdot s_1 \cdot h_{\alpha-\beta}(c^{-1}) \cdot x_{\alpha-\beta}(\epsilon acm').\end{equation}
Notice that $x_{-\beta}(m'X) \in P_\alpha(0)$, therefore from the definition of $S_\alpha(a, -)$ we obtain that
\begin{align*}
 S_\alpha\left(a, x_{-\beta}(m'X) \cdot x_{-\alpha}(mX)\right) \cdot h_{\alpha}^{-1}(1+am)\cdot h_{\beta-\alpha}(c^{-1})\cdot x_{\beta - \alpha}(-bc) \\ 
 = S_\alpha(a, x_{-\beta}(m'X)) \cdot S_\alpha(a, x_{-\alpha}(mX)) \cdot h_\alpha^{-1} (1 + am) \cdot h_{\beta - \alpha}(c^{-1}) \cdot x_{\beta - \alpha}(-bc) &\text{ by~\cref{lem:Smult}}\\
 = S_\alpha(a, x_{-\beta}(m'X)) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot s_0 \cdot h_{\beta - \alpha}(c^{-1}) \cdot x_{\beta - \alpha}(-bc) &\text{ by~\cref{lem:Tulenbaev-formula}}\\
 = x_{-\beta}(m'X) \cdot x_{\alpha - \beta}(\epsilon am') \cdot x_{\beta - \alpha}(-bc^{-1}) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot s_0 \cdot h_{\beta - \alpha}(c^{-1}) &\text{ by~\eqref{Chevalley-CCF1},~\eqref{Chevalley-CCF2}, \eqref{eq:conj-h-x}}\\
 \begin{split}= x_{-\beta}(m'X) \cdot x_{\beta - \alpha}(-b) \cdot s_1 \cdot h_{\alpha-\beta}(c^{-1}) \cdot \hspace{100pt} \\ \cdot x_{\alpha-\beta}\left(\epsilon acm'\right) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot s_0 \cdot h_{\beta - \alpha}(c^{-1}) &\text{ by~\eqref{eq:computation-SR-obtuse}} \end{split}\\
 =  x_{\beta - \alpha}(-b) \cdot x_{-\beta}(m'X) \cdot x_{-\alpha}(-\epsilon bm'X) \cdot s_1 \cdot x_{\alpha-\beta}(\epsilon ac^{-1}m') \cdot x_{-\alpha}\left(\tfrac{cmX}{1+am}\right) \cdot s_0 &\text{ by~\eqref{Chevalley-CCF1},~\eqref{eq:conj-h-x},~\eqref{eq:h-inv}}\\
 = x_{\beta-\alpha}(-b) \cdot x_{-\beta}(m'c^{-1}X) \cdot x_{\alpha-\beta}(\epsilon ac^{-1}m') \cdot s_1 \cdot s_0 \cdot x_{-\alpha}\left(-\epsilon bm' X + \tfrac{cmX}{1+am}\right) &\text{ by~\eqref{Chevalley-CCF1}.}
\end{align*}
Denote by $h_0$ the product of all factors in the above formula except the first and the last one. It is clear that $h_0$ lies in $P_\alpha(0)$. By~\cref{K-a-b} we can decompose $g$ as $g_0 \cdot x_{-\beta}(m'X) \cdot x_{-\alpha}(mX)$ for some $g_0 \in K(\alpha, \beta)$, $m, m' \in M$. The required assertion now follows from Lemmas~\ref{conj-S-a-x-bma}, \ref{lem:Smult} and the above computation:
\begin{multline} \nonumber
 x_{\beta - \alpha}(b) \cdot S_\alpha(a, g) \cdot h_{\alpha}^{-1}(1+am)\cdot h_{\beta - \alpha}(c^{-1})\cdot x_{\beta - \alpha}(-bc) = \\
  = x_{\beta - \alpha}(b) \cdot S_\alpha(a, g_0) \cdot x_{\beta-\alpha}(-b) \cdot h_0 \cdot x_{-\alpha}\left(\tfrac{m -\epsilon bm'}{1+am}\right) \in P_\alpha(\tfrac{m -\epsilon bm'}{1+am}). \qedhere
\end{multline}
\end{proof}

>>>>

Now set $\widetilde{G}_n = G(\rD_n, -) = \mathrm{Spin}(2n, -)$ and $\widetilde{P}_n = \mathrm{E}(\rD_n, -) = \mathrm{Epin}(2n, -)$.
Notice that for $n\geq 3$ the groups
\begin{align*}
   \KO_2(2n, R) &= \K_2^{G_n,P_n} = \Ker(\St(\rD_n, R) \to \mathrm{SO}(2n, R)), \\
 \K_2(\rD_n, R) &= \K_2^{\widetilde{G}_n,\widetilde{P}_n} = \Ker(\St(\rD_n, R) \to \mathrm{Spin}(2n, R))
\end{align*}
are related via the exact sequence
\[ \xymatrix{ 1 \ar[r] & \K_2(\rD_n, R) \ar[r] & \KO_2(2n, R) \ar[r] & \mu_2(R) \ar[r] & 1,} \]
which can be obtained from the exact sequence on p.~189 of~\cite{Ba74} using the snake lemma (cf. also with~\cite[Corollary~4.3.5]{Ba74}).
On the other hand, from~\cref{lem:lowerKgroups} it follows that $\KO_3(2n, R) \cong \mathrm{K}_3^{\widetilde{G}_n, \widetilde{P}_n}$.

@@

\begin{lemma} \label{SR:obtuse} For $g \in j_+(K(A[X], M[X])) \cap P_\alpha(m) \cap P_\beta(m')$ and $a, b\in A$ the element \[ x_{\beta - \alpha}(b) \cdot S_\alpha(a, g) \cdot h_{\alpha}^{-1}(1+am)\cdot h_{\beta - \alpha}((1 + \epsilon abm')^{-1})\cdot x_{\beta - \alpha}(-b(1 + \epsilon abm')) \]
 belongs to $P_\alpha\left(\tfrac{m - \epsilon bm'}{1+am}\right)$, where $\epsilon = N_{\alpha, -\beta}$.
\end{lemma}
\begin{proof} Set $c = 1 + \epsilon a b m'$, $s_0 = \langle a, m \rangle$, $s_1 = \langle \epsilon am', -bc^{-1} \rangle$.
Notice that $1 - \epsilon abc^{-1}m' = c^{-1}$, therefore by~\eqref{eq:dennis-stein}
\begin{equation} \label{eq:computation-SR-obtuse} x_{\alpha-\beta}(\epsilon am') \cdot x_{\beta-\alpha}(-bc^{-1}) = x_{\beta-\alpha}(-b) \cdot s_1 \cdot h_{\alpha-\beta}(c^{-1}) \cdot x_{\alpha-\beta}(\epsilon acm').\end{equation}
Notice that $x_{-\beta}(m'X) \in P_\alpha(0)$, therefore from the definition of $S_\alpha(a, -)$ we obtain that
\begin{align*}
 S_\alpha\left(a, x_{-\beta}(m'X) \cdot x_{-\alpha}(mX)\right) \cdot h_{\alpha}^{-1}(1+am)\cdot h_{\beta-\alpha}(c^{-1})\cdot x_{\beta - \alpha}(-bc) \\ 
 = S_\alpha(a, x_{-\beta}(m'X)) \cdot S_\alpha(a, x_{-\alpha}(mX)) \cdot h_\alpha^{-1} (1 + am) \cdot h_{\beta - \alpha}(c^{-1}) \cdot x_{\beta - \alpha}(-bc) &\text{ by~\cref{lem:Smult}}\\
 = S_\alpha(a, x_{-\beta}(m'X)) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot s_0 \cdot h_{\beta - \alpha}(c^{-1}) \cdot x_{\beta - \alpha}(-bc) &\text{ by~\cref{lem:Tulenbaev-formula}}\\
 = x_{-\beta}(m'X) \cdot x_{\alpha - \beta}(\epsilon am') \cdot x_{\beta - \alpha}(-bc^{-1}) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot s_0 \cdot h_{\beta - \alpha}(c^{-1}) &\text{ by~\eqref{Chevalley-CCF1},~\eqref{Chevalley-CCF2}, \eqref{eq:conj-h-x}}\\
 \begin{split}= x_{-\beta}(m'X) \cdot x_{\beta - \alpha}(-b) \cdot s_1 \cdot h_{\alpha-\beta}(c^{-1}) \cdot \hspace{100pt} \\ \cdot x_{\alpha-\beta}\left(\epsilon acm'\right) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot s_0 \cdot h_{\beta - \alpha}(c^{-1}) &\text{ by~\eqref{eq:computation-SR-obtuse}} \end{split}\\
 =  x_{\beta - \alpha}(-b) \cdot x_{-\beta}(m'X) \cdot x_{-\alpha}(-\epsilon bm'X) \cdot s_1 \cdot x_{\alpha-\beta}(\epsilon ac^{-1}m') \cdot x_{-\alpha}\left(\tfrac{cmX}{1+am}\right) \cdot s_0 &\text{ by~\eqref{Chevalley-CCF1},~\eqref{eq:conj-h-x},~\eqref{eq:h-inv}}\\
 = x_{\beta-\alpha}(-b) \cdot x_{-\beta}(m'c^{-1}X) \cdot x_{\alpha-\beta}(\epsilon ac^{-1}m') \cdot s_1 \cdot s_0 \cdot x_{-\alpha}\left(-\epsilon bm' X + \tfrac{cmX}{1+am}\right) &\text{ by~\eqref{Chevalley-CCF1}.}
\end{align*}
Denote by $h_0$ the product of all factors in the above formula except the first and the last one. It is clear that $h_0$ lies in $P_\alpha(0)$. By~\cref{K-a-b} we can decompose $g$ as $g_0 \cdot x_{-\beta}(m'X) \cdot x_{-\alpha}(mX)$ for some $g_0 \in K(\alpha, \beta)$, $m, m' \in M$. The required assertion now follows from Lemmas~\ref{conj-S-a-x-bma}, \ref{lem:Smult} and the above computation:
\begin{multline} \nonumber
 x_{\beta - \alpha}(b) \cdot S_\alpha(a, g) \cdot h_{\alpha}^{-1}(1+am)\cdot h_{\beta - \alpha}(c^{-1})\cdot x_{\beta - \alpha}(-bc) = \\
  = x_{\beta - \alpha}(b) \cdot S_\alpha(a, g_0) \cdot x_{\beta-\alpha}(-b) \cdot h_0 \cdot x_{-\alpha}\left(\tfrac{(m -\epsilon bm')X}{1+am}\right) \in P_\alpha(\tfrac{m -\epsilon bm'}{1+am}). \qedhere
\end{multline}
\end{proof}
